We first begin by importing some useful functions and libraries.
```{r}
library(tidyverse)
library(reshape2)
library(latex2exp)

setwd("/Users/mathieu/Lab/company-electricity-statistical-analysis/src/")
source("utils.r")
```

We remember that the daily electricity consumption (in MWh) is modelled as a random variable $X$ with the following probability density function (PDF), 
\begin{\begin{equation}
  f_{\theta_1, \theta_2}(x) = 
  \begin{cases}
    \frac{\theta_1 \theta_2^{\theta_1}}{x^{\theta_1 + 1}}, &\quad x \geq \theta_2 \\
    0,                                                     &\quad \text{otherwise}
  \end{cases}
(\#eq:PDF)
\end{equation}}

Setting, $\theta_1^0 = 3$ and $\theta_2^0 = 1$, we want to generate and i.i.d. sample of size $n=20$ from the density $f_{\theta_1^0, \theta_2^0}$ making use of **inverse transform sampling**. 
We then use this sample to compute $\hat{G}_{\text{MLE}}$ and $\hat{G}_{\text{MME}}$.

```{r}
# set a seed for reproductability
set.seed(42)

# Generate sample of size n = 20 by using inverse transform sampling
rv_vector <- inverse_transform_sampling(n = 20, inv_cdf = inv_cdf)

# plot an histogram of the random variable vector
par(mfrow = c(1,1))
hist(rv_vector, breaks = 50, freq = FALSE, xlab = "X", main = "random sample")

# compute Gini coefficients
gini_mle(rv_vector = rv_vector, n = 20)
gini_mme(rv_vector = rv_vector, n = 20)
```

We then want to repeat the data generating process for $N = 1000$ times (still with same sample size and same parameters for $\theta$).

We get a sample of size $N = 1000$ of these estimators of the Gini coefficient.
We can make a **histogram** and a **boxplot** of these two samples.
```{r}
# set a seed for reproductability
set.seed(42)

# Generate N = 1000 times the sample
x <- sim(N = 1000, n = 20, f = inverse_transform_sampling, inv_cdf)

gini_mle_sample <- x["gini-mle-sample", ]
gini_mme_sample <- x["gini-mme-sample", ]

# histogram of gini samples
par(mfrow = c(1, 2))
hist(gini_mle_sample, breaks = 50, main = "", xlab = TeX(r"($\hat{G}_{\textrm{MLE}}$)"), col = "darkorange")
abline(v = gini_theoretical(theta_1), col = "darkgreen", lty = 2)
hist(gini_mme_sample, breaks = 50, main = "", xlab = TeX(r"($\hat{G}_{\textrm{MME}}$)"), col = "steelblue")
abline(v = gini_theoretical(theta_1), col = "darkgreen", lty = 2)
legend("topright", c("Gini MLE", "Gini MME", "True Gini"), fill = c("darkorange", "steelblue", "darkgreen"))

# boxplot of gini samples
par(mfrow = c(1, 2))
boxplot(gini_mle_sample, col = "grey", ylab = TeX(r"($\hat{G}_{\textrm{MLE}}$)"))
abline(h = gini_theoretical(theta_1), col = "darkgreen", lty = 2)
boxplot(gini_mme_sample, col = "grey", ylab = TeX(r"($\hat{G}_{\textrm{MME}}$)"))
abline(h = gini_theoretical(theta_1), col = "darkgreen", lty = 2)
```

We use these samples to estimate the **bias**, the **variance** and the **mean squared error (MSE)** of both estimators.
```{r}
bias_mle <- bias(estimator_sample = gini_mle_sample, theoretical_value = gini_theoretical(theta_1))
bias_mme <- bias(estimator_sample = gini_mme_sample, theoretical_value = gini_theoretical(theta_1))

variance_mle <- var(gini_mle_sample)
variance_mme <- var(gini_mme_sample)

mse_mle <- mse(estimator_sample = gini_mle_sample, theoretical_value = gini_theoretical(theta_1))
mse_mme <- mse(estimator_sample = gini_mme_sample, theoretical_value = gini_theoretical(theta_1))

print(bias_mle)
print(variance_mle)
print(mse_mle)

print(bias_mme)
print(variance_mme)
print(mse_mme)
```

We repeat the simulation $N = 1000$ times to generate multiple samples of the gini estimators based on a sample of sizes $n = 20, 40, 60, 80, 100, 150, 200, 300, 400, 500$ generated from the density function using inverse transform sampling.

We can then compare the the **bias**, the **variance** and the **mean squared error (MSE)** of both estimators for the different sample sizes via a separate plot for each quantity as a function of $n$.
```{r}
# set a seed for reproductability
set.seed(42)

sample_sizes <- c(20, 40, 60, 80, 100, 150, 200, 300, 400, 500) 

statistical_quantities <- c("gini-bias-mle", "gini-bias-mme", "gini-variance-mle", "gini-variance-mme", "gini-mse-mle", "gini-mse-mme")

matrix <- matrix(NA, nrow = length(statistical_quantities), ncol = length(sample_sizes))
rownames(matrix) <- statistical_quantities
colnames(matrix) <- sample_sizes

# create the different samples, one for each sample size n
for (n in sample_sizes) {
  x <- sim(N = 1000, n = n, f = inverse_transform_sampling, inv_cdf)

  gini_mle_sample <- x["gini-mle-sample", ]
  gini_mme_sample <- x["gini-mme-sample", ]

  bias_mle <- bias(estimator_sample = gini_mle_sample, theoretical_value = gini_theoretical(theta_1))
  bias_mme <- bias(estimator_sample = gini_mme_sample, theoretical_value = gini_theoretical(theta_1))

  variance_mle <- var(gini_mle_sample)
  variance_mme <- var(gini_mme_sample)

  mse_mle <- mse(estimator_sample = gini_mle_sample, theoretical_value = gini_theoretical(theta_1))
  mse_mme <- mse(estimator_sample = gini_mme_sample, theoretical_value = gini_theoretical(theta_1))

  matrix[, as.character(n)] <- c(bias_mle, bias_mme, variance_mle, variance_mme, mse_mle, mse_mme)
}

# create dataframes
bias_df <- data.frame(n = sample_sizes, mle = matrix["gini-bias-mle", ], mme = matrix["gini-bias-mme", ])
var_df <- data.frame(n = sample_sizes, mle = matrix["gini-variance-mle", ], mme = matrix["gini-variance-mme", ])
mse_df <- data.frame(n = sample_sizes, mle = matrix["gini-mse-mle", ], mme = matrix["gini-mse-mme", ])

# converting to long format for easier plotting
bias_df <- melt(bias_df, id = "n")
var_df <- melt(var_df, id = "n")
mse_df <- melt(mse_df, id = "n")

par(mfrow = c(1,1))

# we compare the gini estimators through a plot as a function of n
ggplot(bias_df, aes(x = n, y = value, color = variable)) +
  geom_point() + 
  geom_line() +
  labs(x = "n", y = "bias") +
  scale_fill_hue(labels = c("Gini MLE bias", "Gini MME bias")) +
  theme(legend.title = element_blank())

ggplot(var_df, aes(x = n, y = value, color = variable)) +
  geom_point() + 
  geom_line() +
  labs(x = "n", y = "variance") +
  scale_fill_hue(labels = c("Gini MLE: variance", "Gini MME: variance")) +
  theme(legend.title = element_blank())

ggplot(mse_df, aes(x = n, y = value, color = variable)) +
  geom_point() + 
  geom_line() +
  labs(x = "n", y = "mean squared error") + 
  scale_fill_hue(labels = c("Gini MLE: mse", "Gini MME: mse")) +
  theme(legend.title = element_blank())
```

Finally, we create an histogram of $\sqrt{n}(\hat{G}_{\text{MLE}} - G_{\theta_1^0, \theta_2^0})$ for $n = 20$, $n = 100$ and $n = 100$ and compare them.
```{r}
sim_n20 <- sim(N = 1000, n = 20, f = inverse_transform_sampling, inv_cdf)
result_n20 <- sqrt(20) * (sim_n20["gini-mle-sample", ] - gini_theoretical(theta_1))

sim_n100 <- sim(N = 1000, n = 100, f = inverse_transform_sampling, inv_cdf)
result_n100 <- sqrt(100) * (sim_n100["gini-mle-sample", ] - gini_theoretical(theta_1))

sim_n500 <- sim(N = 1000, n = 500, f = inverse_transform_sampling, inv_cdf)
result_n500 <- sqrt(500) * (sim_n500["gini-mle-sample", ] - gini_theoretical(theta_1))

par(mfrow = c(1, 3))
hist(result_n20, breaks = 50, main = "sample size: n = 20", xlab = TeX(r"($\sqrt{n}(\hat{G}_{\textrm{MLE}} - G_{\theta_1^0, \theta_2^0})$)"), col = "coral")
hist(result_n100, breaks = 50, main = "sample size: n = 100", xlab = TeX(r"($\sqrt{n}(\hat{G}_{\textrm{MLE}} - G_{\theta_1^0, \theta_2^0})$)"), col = "aquamarine3")
hist(result_n500, breaks = 50, main = "sample size: n = 500", xlab = TeX(r"($\sqrt{n}(\hat{G}_{\textrm{MLE}} - G_{\theta_1^0, \theta_2^0})$)"), col = "steelblue")
```